---
title: "visualization"
author: "Nelson Gaillard"
date: "2025-09-25"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

Import the weather data

```{r}
data("weather_df")
```

Making our first plot :-)

```{r}
ggplot(weather_df, mapping = aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax))+
  geom_point()
```

```{r}
ggp_weather_scatterplot = 
  weather_df |>
    ggplot(aes(x = tmin, y = tmax))+
    geom_point()

ggp_weather_scatterplot
```

```{r}
weather_df |>
  filter(is.na(tmax))
```


## Fancier scatterplots!

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)
```

Where you define aestherics can matter

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE, size = 0.6)
```

use faceting real quick

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE, size = 0.6) +
  facet_grid(. ~ name)

weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE, size = 0.6) +
  facet_grid(name ~ .)
```

Let's make a somewhat more interesting scatterplot

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE, size = 0.6) +
  facet_grid(. ~ name)
```

Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperatures, and overlays a linear regression line (using options in geom_smooth())

```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax_fahr = tmax * (9/5) + 32,
    tmin_fahr = tmin * (9/5) + 32
  ) |>
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
    geom_point() +
    geom_smooth(se = FALSE, method = "lm")
```

## Small things

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)

weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_smooth(se = FALSE) +
  geom_point()
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "purple")
```

## Univariate plots

```{r}
weather_df |>
  ggplot(aes(x = tmin)) +
  geom_histogram(color = "white", fill = "red")
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(name ~ .)
```

Maybe a density plot?

```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(lpha = 0.2)
```

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot(aes(fill = name))
```

violin plots

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()
```

ridge plot

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = name, fill = name)) +
  geom_density_ridges()
```

Learning assessment: make plots that compare precipitation across locations.

```{r}
weather_df |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density()

weather_df |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_histogram() +
  facet_grid(name ~ .)

weather_df |>
  filter(prcp > 5, prcp < 1000) |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.2)
```



