---
title: "visualization"
author: "Nelson Gaillard"
date: "2025-09-25"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
library(patchwork)
```

## Visualization 1

Import the weather data

```{r}
data("weather_df")
```

Making our first plot :-)

```{r}
ggplot(weather_df, mapping = aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax))+
  geom_point()
```

```{r}
ggp_weather_scatterplot = 
  weather_df |>
    ggplot(aes(x = tmin, y = tmax))+
    geom_point()

ggp_weather_scatterplot
```

```{r}
weather_df |>
  filter(is.na(tmax))
```


## Fancier scatterplots!

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)
```

Where you define aestherics can matter

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE, size = 0.6)
```

use faceting real quick

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE, size = 0.6) +
  facet_grid(. ~ name)

weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE, size = 0.6) +
  facet_grid(name ~ .)
```

Let's make a somewhat more interesting scatterplot

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE, size = 0.6) +
  facet_grid(. ~ name)
```

Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperatures, and overlays a linear regression line (using options in geom_smooth())

```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax_fahr = tmax * (9/5) + 32,
    tmin_fahr = tmin * (9/5) + 32
  ) |>
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
    geom_point() +
    geom_smooth(se = FALSE, method = "lm")
```

## Small things

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)

weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_smooth(se = FALSE) +
  geom_point()
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "purple")
```

## Univariate plots

```{r}
weather_df |>
  ggplot(aes(x = tmin)) +
  geom_histogram(color = "white", fill = "red")
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(name ~ .)
```

Maybe a density plot?

```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(lpha = 0.2)
```

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot(aes(fill = name))
```

violin plots

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()
```

ridge plot

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = name, fill = name)) +
  geom_density_ridges()
```

Learning assessment: make plots that compare precipitation across locations.

```{r}
weather_df |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density()

weather_df |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_histogram() +
  facet_grid(name ~ .)

weather_df |>
  filter(prcp > 5, prcp < 1000) |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.2)
```

## Saving and embedding plots

saving plots

```{r}
ggp_weather_violin =
  weather_df |>
    ggplot(aes(x = name, y = tmin, fill = name)) +
    geom_violin()
  
ggp_weather_violin

ggsave("visualization_files/violin_plot.pdf", ggp_weather_violin,
       width = 8, height = 6)
```

embedding plots

```{r, fig.width = 6, fig.height = 4}



```

## Visualization 2

```{r}
data("weather_df")
```

Let's make our basic scatterplot

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp", 
    title = "temperature scatterplot",
    caption = "Data from NOAA",
    color = "location"
  )
```

#### Scales

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp", 
    title = "temperature scatterplot",
    caption = "Data from NOAA",
    color = "location"
  ) +
  scale_x_continuous(
    breaks = c(-20, 0, 25),
    labels = c("-20C", "0", "25")
  ) +
  scale_y_continuous(
    trans = "sqrt",
    limits = c(10, 30)
  ) +
  viridis::scale_color_viridis(
    discrete = TRUE
  )
```

### Themes

Make my base plot

```{r}
ggplot_temperature = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5) +
  labs(
    x = "Minimum daily temp",
    y = "Maximum daily temp", 
    title = "temperature scatterplot",
    caption = "Data from NOAA",
    color = "location"
  ) +
  scale_x_continuous(
    breaks = c(-20, 0, 25),
    labels = c("-20C", "0", "25")
  ) +
  scale_y_continuous(
    trans = "sqrt",
    limits = c(10, 30)
  ) +
  viridis::scale_color_viridis(
    discrete = TRUE
  )
```

Update my base plot

```{r}
ggplot_temperature +
  theme_bw() +
  theme(legend.position = "bottom")
```

### Adding data into geoms

```{r}
central_park_df = 
  weather_df |>
  filter(name == "CentralPark_NY")

molokai_df =
  weather_df |>
  filter(name == "Molokai_HI")

ggplot(data = molokai_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```

### `patchwork`

Make three plots and combine using patchwork.

```{r}
ggp_tmax_tmin =
  weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "none")

ggp_prec_density =
  weather_df |>
  filter(prcp > 0) |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "none")

ggp_temp_season = 
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

(ggp_tmax_tmin + ggp_prec_density)

```

## Data manipulation

Let's make temperature violin plots.

```{r}
weather_df |>
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhold_WA"))) |>
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

```{r}
weather_df |>
  mutate(name = fct_reorder(name, tmax)) |>
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

What about tidiness?

```{r}
pulse_df = 
  haven::read_sas("data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi"
  ) |>
  mutate(visit = fct_inorder(visit))

pulse_df |>
  ggplot(aes(x = visit, y = bdi)) +
  geom_boxplot()
```

Making that plot that Jeff is asking us to make

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv", na = c("NA", ".", ""), skip = 3) |>
  janitor::clean_names() |>
  mutate(
    sex = case_match(
      sex, 
      1 ~ "male",
      2 ~ "female"
    )
  )

litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "tx_day"), sep = 3)
  
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number")

fas_df |>
  select(pd_ears:tx_day) |>
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    names_prefix = "pd_",
    values_to = "pn_day") |>
  mutate(outcome = fct_reorder(outcome, pn_day)) |>
  drop_na() |>
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
  facet_grid(tx_day ~ outcome)
  
```


















